Compararison of analytical and artificial neural network models in predicting hot deformation behavior of AISI P20 and interpolation-extrapolation capacity validation

### 2022-10-19 Article.tex
Update of Johnson-Cook parameters and script

% !TeX encoding = UTF-8
\documentclass[twoside,english,1p,final,sort&compress]{elsarticle}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\pagestyle{headings}
\usepackage{amsthm}
%\usepackage{gensymb}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{accents}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{subcaption}
\usepackage{adjustbox}
\usepackage{placeins}
\usepackage[colorlinks = true]{hyperref}

@@

\begin{abstract}
The high-temperature deformation behavior of the AISI P20 alloy was investigated by performing hot compression tests over a wide range of strains ($0.025 - 0.7$),  strain rates ($0.001\ \text{s}^{-1} - 5\ \text{s}^{-1}$) and temperatures ($1050\text{°C}-1250\text{°C}$). Based on the experimental data (stress-strain curves), empirical models as the Johnson-Cook (JC), Modified-Zerilli-Armstrong (MZA), Hansel-Spittel (HS), Arhenius (AR) and PTM on the one hand and an ANN model on the other hand were developed to predict the flow behavior of this alloy. Then, a comparative study between the analytical models and the ANN model was carried out on the basis of the calculation of some coefficients allowing to test the efficiency of each model. These coefficients are the correlation coefficient $R$, the Average Absolute  Relative Error $\AARE$ and the Root Mean Square Error ($\RMSE$). From these parameters it was found that the JC, MZA and HS models are not suitable for predicting the behavior of AISI P20 alloy whereas the AR and PTM models can be used for most strain rates to predict the flow behavior of this alloy. For the strain rate ($\mdot\varepsilon^p = 0.01\ \text{s}^{-1}$)  the AR model fails to describe the flow for temperatures $1050$°C and $1100$°C  while in the case of the PTM model, the strain rate  ($\mdot\varepsilon^p = 0.1\ \text{s}^{-1}$) is not well described. This is due to the fact that this model is a sequence of polynomials which can create oscillations if the degree is high. Regarding the ANN model, the architecture retained is composed of two hidden layers of which the first hidden layer contains $15$ neurons and the second contains $7$ neurons and the input layer contains $3$ neurons corresponding to the strain ($\varepsilon^p$), strain rate ($\mdot\varepsilon^p$) and temperature ($T$) and the output layer contains a single neuron corresponding to the flow stress ($\sigma^y$). The results of the ANN model show a very good agreement between the predictions and the experiment. This demonstrates that this  formulation of constitutive law exceeds highly the analytical models in terms of prediction. To validate the performance of these models their capacities to interpolate and extrapolate the data were also tested where all the models relatively prove a good interpolation capacity whereas this is not the case for extrapolation where only the AR, PTM and ANN models are acceptable but still mentioning that the ANN model remains the best in both interpolation and extrapolation.
\end{abstract}

@@

The material used in this study is the modified carbon alloy AISI P20 whose chemical composition is given in Table \ref{tab:Composition}. Cylindrical samples were machined with an initial diameter of $10$~mm and a height of $15$~mm. A batch of $30$ hot compression tests were performed on a Gleeble-3800 thermomechanical simulator (see Figure \ref{fig:Gleeble3800}), for the five temperature levels of $1050$°C, $1100$°C, $1150$°C, $1200$°C and $1250$°C with the six strain rates of $0.001~\text{s}^{-1}$, $0.01~\text{s}^{-1}$, $0.1~\text{s}^{-1}$, $1.0~\text{s}^{-1}$, $2.0~\text{s}^{-1}$ and $5.0~\text{s}^{-1}$.

@@

Thin tantalum sheets were used as a lubricating material at the contact surface of the anvils and samples to minimize friction during testing. As shown in Figure \ref{fig:GleebleProcess}, the samples were heated to a temperature of $1260$°C with a heating rate of $2$°C/s and held at this temperature for $5$~min to eliminate thermal gradients. They were then cooled with a rate of $1$°C/s to the test temperature and then held at constant temperature for $1$~min before deformation. During the compression phase, the temperature of the specimen is kept constant by the thermal control system of the machine. After compression, the specimen is quickly quenched to freeze its microstructure for later analysis. Figure \ref{fig:GleebleProcess} also shows the aspects of the specimens before and after the compression test: $h_0$ and $R_0$ are the height and radius before compression and $h$, $R_M$ and $R_T$ are the height, large radius and small radius of the specimen after compression, respectively.

\begin{figure}[!ht]
\centering
\includegraphics[width=0.8\columnwidth]{Figures/GleebleProcess}
\caption{Schematic diagram of the experimental process.}
\label{fig:GleebleProcess}
\end{figure}

%------------------------------------------------------------------------
\subsection{Compression tests results}
%------------------------------------------------------------------------

The set of flow stress $\sigma^y$ versus plastic strain $\varepsilon^p$ curves obtained from compression tests performed on the Gleeble-3800 simulator for each test condition ($6$ strain rates and $5$ temperatures) is presented in Figure \ref{fig:rawData}. The overall behavior of these curves shows that the flow stress $\sigma^y$ increases with increasing strain rate $\mdot\varepsilon^p$ but decreases with increasing temperature $T$. It should be noted that the plastic strain also influences the flow stress. Indeed, for the lowest strain rates $\mdot\varepsilon^p$, the flow stress $\sigma^y$ increases with the plastic strain $\varepsilon^p$ until a value of about $\varepsilon^p=0.2$ to $0.3$, then decreases to maintain a more or less constant value until the end of the test. For the highest strain rates ($1.0~\text{s}^{-1}$, $2.0~\text{s}^{-1}$, and $5.0~\text{s}^{-1}$), the flow stress increases throughout the test. One reason for the slight increase in stress at low strain rates when the strain is large is due to friction between the sample and the anvil during the test. For low strain rates, the effect of lubrication decreases over time and thus friction increases. This results in a progressive increase of the flow stress.

@@

To explain the overall flow stress behavior observed in this test, we will rely on two fundamental phenomena that can be observed during a compression test: Dynamic Recovery (DRC) and Dynamic Recrystallization (DRX). At the beginning of the deformation of the sample in compression, the amount of dislocations increases and they begin to move within the material, which leads to a rapid increase in stress (strain hardening phenomenon). As the strain continues to increase, many dislocations are in motion, leading to a stress relaxation phenomenon, generally known as DRC. This DRC mechanism contributes to the decrease in the number of dislocations due to their combination with another dislocation or due to their integration into a grain boundary. Some of these dislocations reorganize and then form sub-grains within the initial grains of the material. The combination of these dislocations moderates the increase in the number of dislocations related to work hardening and thus controls the amount of strain energy stored in the material, which influences the increase in flow stress. The dynamic recrystallization concerns the germination of new grains, at the borders or inside the initial grains. These two phenomena explain the different shapes of the flow stress curves observed for this alloy. As an example, Figure \ref{fig:Micrography} show the microstructures of AISI P20 held during $5$~min at $1260$°C before and after the compression phase. As observed on the flow curves, it can be seen by comparing these two micrographs that recrystallization is almost complete for this temperature and strain rate within the sample \OP{What are the temperature and strain rate related with this micrography}. From the results reported in Figure \ref{fig:rawData}, the identification of the parameters of the six flow laws proposed in this work will be presented in the next section in order to predict the behavior of the material in any test condition.

@@

\begin{equation}
\sigma^y=\left(A+B\varepsilon^{p^{n}}\right) \left[1+C\ln\left(\frac{\mdot\varepsilon^p}{\mdot\varepsilon_0}\right)\right] \left[1-\left(\frac{T-T_0}{T_m-T_0}\right)^{m}\right],\label{eq:Johnson-Cook}
\end{equation}
where $\sigma^y$ is the flow stress, $A$ is the initial elastic limit of the material, $B$ is the strain hardening coefficient, $n$ is the strain hardening exponent, $C$ and $m$ are the material constants that describe the strain rate hardening coefficient and the thermal softening coefficient, respectively. The other values are reference ones, so that $\mdot\varepsilon_0$ is the reference strain rate, $T_m$ and $T_0$ are the melting temperature ($1460$°C in our case) and the reference temperature, respectively.

For the determination of the parameters of the analytical models, the reference values for strain rate and temperature are $\mdot\varepsilon_0=0.001~\text{s}^{-1}$ and $T_0=1050$°C respectively.

The procedure used for the determination of the parameters of the Johnson-Cook law is in accordance with the one proposed by He \eal \cite{He-2013}. This method allows obtaining one after the other the $5$ parameters in the order $A$, $B$, $n$, $C$ and $m$. Thus, according to the experimental data, the initial elastic limit of the material at the reference strain rate and reference temperature is $A=11.76$~MPa. For the determination of the constants $B$ and $n$, from the results of the compression test at $T_0$ and $\mdot\varepsilon_0$, it is then possible to determine these two constants by considering only the first term $\left(A+B\varepsilon^{p^{n}}\right)$ in Equation (\ref{eq:Johnson-Cook}). Thus, here, $B=23.51$~MPa and $n=0.066$. Once the parameters $A$, $B$ and $n$ are known, the determination of $C$, considering all the curves at $T=T_0$ then gives $C=0.352$. Finally, the last parameter $m$ is identified from the curves at $\mdot\varepsilon=\mdot\varepsilon_0$ and from the knowledge of the previous and gives $m=0.696$.
All parameters of the Johnson-Cook model are reported in Table \ref{tab:JCparams}.

\begin{table}[h!]
\centering{}
\caption{Parameters' values of the Johnson-Cook flow law for the P20 steel.}
\begin{tabular}{lccccc}
\hline
Parameters & $A~(\text{MPa})$ & $B~(\text{MPa})$ & $n$ & $C$ & $m$\\ \hline
Values & $43.0988$ & $114.518$ &  $0.313969$ & $0.0995438$ & $0.898932$\\ \hline
\label{tab:JCparams}
\end{tabular}
\end{table}

@@

Experiments were conducted on modified AISI P20 carbon alloy to study the applicability and predictive accuracy of five analytical models and one neural network model over a range of temperatures ($1050$°C - $1250$°C, strains ($0.025$ - $0.7$), strain rates ($0.001\ \text{s}^{-1}$ - $5\ \text{s}^{-1}$) and the following conclusions were drawn.

>>>>

% !TeX encoding = UTF-8
\documentclass[twoside,english,1p,final,sort&compress]{elsarticle}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\pagestyle{headings}
\usepackage{amsthm}
\usepackage{gensymb}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{accents}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{subcaption}
\usepackage{adjustbox}
\usepackage{placeins}
\usepackage[colorlinks = true]{hyperref}

@@

\begin{abstract}
The high-temperature deformation behavior of the AISI P20 alloy was investigated by performing hot compression tests over a wide range of strains ($0.025 - 0.7$),  strain rates ($0.001\ \text{s}^{-1} - 5\ \text{s}^{-1}$) and temperatures ($1050\celsius-1250\celsius$). Based on the experimental data (stress-strain curves), empirical models as the Johnson-Cook (JC), Modified-Zerilli-Armstrong (MZA), Hansel-Spittel (HS), Arhenius (AR) and PTM on the one hand and an ANN model on the other hand were developed to predict the flow behavior of this alloy. Then, a comparative study between the analytical models and the ANN model was carried out on the basis of the calculation of some coefficients allowing to test the efficiency of each model. These coefficients are the correlation coefficient $R$, the Average Absolute  Relative Error $\AARE$ and the Root Mean Square Error ($\RMSE$). From these parameters it was found that the JC, MZA and HS models are not suitable for predicting the behavior of AISI P20 alloy whereas the AR and PTM models can be used for most strain rates to predict the flow behavior of this alloy. For the strain rate ($\mdot\varepsilon^p = 0.01\ \text{s}^{-1}$)  the AR model fails to describe the flow for temperatures $1050\celsius$ and $1100\celsius$  while in the case of the PTM model, the strain rate  ($\mdot\varepsilon^p = 0.1\ \text{s}^{-1}$) is not well described. This is due to the fact that this model is a sequence of polynomials which can create oscillations if the degree is high. Regarding the ANN model, the architecture retained is composed of two hidden layers of which the first hidden layer contains $15$ neurons and the second contains $7$ neurons and the input layer contains $3$ neurons corresponding to the strain ($\varepsilon^p$), strain rate ($\mdot\varepsilon^p$) and temperature ($T$) and the output layer contains a single neuron corresponding to the flow stress ($\sigma^y$). The results of the ANN model show a very good agreement between the predictions and the experiment. This demonstrates that this  formulation of constitutive law exceeds highly the analytical models in terms of prediction. To validate the performance of these models their capacities to interpolate and extrapolate the data were also tested where all the models relatively prove a good interpolation capacity whereas this is not the case for extrapolation where only the AR, PTM and ANN models are acceptable but still mentioning that the ANN model remains the best in both interpolation and extrapolation.
\end{abstract}

@@

The material used in this study is the modified carbon alloy AISI P20 whose chemical composition is given in Table \ref{tab:Composition}.
Cylindrical samples were machined with an initial diameter of $10$~mm and a height of $15$~mm.
A batch of $30$ hot compression tests were performed on a Gleeble-3800 thermomechanical simulator (see Figure \ref{fig:Gleeble3800}), for the five temperature levels of $1050\celsius$, $1100\celsius$, $1150\celsius$, $1200\celsius$ and $1250\celsius$ with the six strain rates of $0.001~\text{s}^{-1}$, $0.01~\text{s}^{-1}$, $0.1~\text{s}^{-1}$, $1.0~\text{s}^{-1}$, $2.0~\text{s}^{-1}$ and $5.0~\text{s}^{-1}$.

@@

Thin tantalum sheets were used as a lubricating material at the contact surface of the anvils and samples to minimize friction during testing.
As shown in Figure \ref{fig:GleebleProcess}, the samples were heated to a temperature of $1260\celsius$ with a heating rate of $2\celsius$/s and held at this temperature for $5$~min to eliminate thermal gradients.
They were then cooled with a rate of $1\celsius$/s to the test temperature and then held at constant temperature for $1$~min before deformation.
During the compression phase, the temperature of the specimen is kept constant by the thermal control system of the machine.
After compression, the specimen is quickly quenched to freeze its microstructure for later analysis.
Figure \ref{fig:GleebleProcess} also shows the aspects of the specimens before and after the compression test: $h_0$ and $R_0$ are the height and radius before compression and $h$, $R_M$ and $R_T$ are the height, large radius and small radius of the specimen after compression, respectively.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.8\columnwidth]{Figures/GleebleProcess}
\caption{Schematic diagram of the experimental process.}
\label{fig:GleebleProcess}
\end{figure}

\OP{Here, a part concerning the treatment of the Gleeble data is missing. You introduce the notions of $h_0$, $R_0$, $h$, $R_M$ and $R_T$ and we have no idea what they are for. How are the Gleeble data transformed into curves, what is the use of the previous values, what are the corrections to be made to the data,...}
%------------------------------------------------------------------------
\subsection{Compression tests results}
%------------------------------------------------------------------------

The set of flow stress $\sigma^y$ versus plastic strain $\varepsilon^p$ curves obtained from compression tests performed on the Gleeble-3800 simulator for each test condition ($6$ strain rates and $5$ temperatures) is presented in Figure \ref{fig:rawData}.
\OP{It is necessary to give indications concerning the structure of the data: how many points, what amplitude of deformation,...} The overall behavior of these curves shows that the flow stress $\sigma^y$ increases with increasing strain rate $\mdot\varepsilon^p$ but decreases with increasing temperature $T$.
It should be noted that the plastic strain also influences the flow stress.
Indeed, for the lowest strain rates $\mdot\varepsilon^p$, the flow stress $\sigma^y$ increases with the plastic strain $\varepsilon^p$ until a value of about $\varepsilon^p=0.2$ to $0.3$, then decreases to maintain a more or less constant value until the end of the test.
For the highest strain rates ($1.0~\text{s}^{-1}$, $2.0~\text{s}^{-1}$, and $5.0~\text{s}^{-1}$), the flow stress increases throughout the test.
One reason for the slight increase in stress at low strain rates when the strain is large is due to friction between the sample and the anvil during the test.
For low strain rates, the effect of lubrication decreases over time and thus friction increases.
This results in a progressive increase of the flow stress.

@@

To explain the overall flow stress behavior observed in this test, we will rely on two fundamental phenomena that can be observed during a compression test: Dynamic Recovery (DRC) and Dynamic Recrystallization (DRX).
At the beginning of the deformation of the sample in compression, the amount of dislocations increases and they begin to move within the material, which leads to a rapid increase in stress (strain hardening phenomenon).
As the strain continues to increase, many dislocations are in motion, leading to a stress relaxation phenomenon, generally known as DRC.
This DRC mechanism contributes to the decrease in the number of dislocations due to their combination with another dislocation or due to their integration into a grain boundary.
Some of these dislocations reorganize and then form sub-grains within the initial grains of the material.
The combination of these dislocations moderates the increase in the number of dislocations related to work hardening and thus controls the amount of strain energy stored in the material, which influences the increase in flow stress.
The dynamic recrystallization concerns the germination of new grains, at the borders or inside the initial grains.
These two phenomena explain the different shapes of the flow stress curves observed for this alloy.
As an example, Figure \ref{fig:Micrography} show the microstructures of AISI P20 held during $5$~min at $1260\celsius$ before and after the compression phase.
As observed on the flow curves, it can be seen by comparing these two micrographs that recrystallization is almost complete for this temperature and strain rate within the sample.
\OP{What are the temperature and strain rate related with this micrography}
From the results reported in Figure \ref{fig:rawData}, the identification of the parameters of the six flow laws proposed in this work will be presented in the next section in order to predict the behavior of the material in any test condition.

@@

\begin{equation}
\sigma^y=\left(A+B\varepsilon^{p^{n}}\right) \left[1+C\ln\left(\frac{\mdot\varepsilon^p}{\mdot\varepsilon_0}\right)\right] \left[1-\left(\frac{T-T_0}{T_m-T_0}\right)^{m}\right],\label{eq:Johnson-Cook}
\end{equation}
where $\sigma^y$ is the flow stress, $A$ is the initial elastic limit of the material, $B$ is the strain hardening coefficient, $n$ is the strain hardening exponent, $C$ and $m$ are the material constants that describe the strain rate hardening coefficient and the thermal softening coefficient, respectively.
The other values are reference ones, so that $\mdot\varepsilon_0$ is the reference strain rate, $T_m$ and $T_0$ are the melting temperature ($1460\celsius$ in our case) and the reference temperature, respectively.

For the determination of the parameters of the analytical models, the reference values for strain rate and temperature are $\mdot\varepsilon_0=0.001~\text{s}^{-1}$ and $T_0=1050\celsius$ respectively.

The procedure used for the determination of the parameters of the Johnson-Cook law is in accordance with the one proposed by He \eal \cite{He-2013}.
This method allows obtaining one after the other the $5$ parameters in the order $A$, $B$, $n$, $C$ and $m$.
Thus, according to the experimental data, the initial elastic limit of the material at the reference strain rate and reference temperature is $A=(11.758$~MPa.
For the determination of the constants $B$ and $n$, from the results of the compression test at $T_0$ and $\mdot\varepsilon_0$, it is then possible to determine these two constants by considering only the first term $\left(A+B\varepsilon^{p^{n}}\right)$ in Equation (\ref{eq:Johnson-Cook}).
Thus, here, $B=25.563$~MPa and $n=0.139$.
Once the parameters $A$, $B$ and $n$ are known, the determination of $C$, considering all the curves at $T=T_0$ then gives $C=0.338$.
Finally, the last parameter $m$ is identified from the curves at $\mdot\varepsilon=\mdot\varepsilon_0$ and from the knowledge of the previous and gives $m=0.730$.
All parameters of the Johnson-Cook model are reported in Table \ref{tab:JCparams}.

\begin{table}[h!]
\centering{}
\caption{Parameters' values of the Johnson-Cook flow law for the P20 steel.}
\begin{tabular}{lccccc}
\hline
Parameters & $A~(\text{MPa})$ & $B~(\text{MPa})$ & $n$ & $C$ & $m$\\ \hline
Values & $11.758$ & $25.563$ &  $0.139$ & $0.338$ & $0.730$\\ \hline
\label{tab:JCparams}
\end{tabular}
\end{table}

@@

Experiments were conducted on modified AISI P20 carbon alloy to study the applicability and predictive accuracy of five analytical models and one neural network model over a range of temperatures ($1050\celsius$ - $1250\celsius$, strains ($0.025$ - $0.7$), strain rates ($0.001\ \text{s}^{-1}$ - $5\ \text{s}^{-1}$) and the following conclusions were drawn.
